%
% file: dedication.tex
% author: Eshwen Bhal
% description: Contains the text for thesis dedication
%

\chapter*{Dedication and acknowledgements}
\begin{SingleSpace}
\initial{T}his work is dedicated to my grandfather, Dato' Mahindar Singh Bhal, who was able to begin this journey with me but sadly unable to finish it. There are far too many people and too little space to individually thank everyone who has accompanied me during this PhD, but I'll try my best.

Firstly, to my supervisor, Dr. Henning Fl\"{a}cher. Your advice and guidance over the course of this degree has been instrumental to achieving it, as well as encouraging my growth as a researcher.

Secondly, to all my colleagues at the University of Bristol. Very little would have been achieved without your help. From scientific discussions in the pub to pub discussions in the physics building, they have all been fruitful either directly or by helping me detach from work. Special thanks are in order to Simone and Sam Maddrell-Mander in my cohort, who have put up with my complaining during the stressful times.

To my friends from Monmouth that include my best friends in the world, Mike, James, Sneddon, (Mini) Sam, and Matt Bristow. You've been on this adventure with me since our school days and had my back the entire time. We've been through the highest, lowest, and most hilarious times together. I can definitively say I would not be the person I am today without you.

To my friends from LTA and those based at CERN, especially Matt Heath, Dwayne and Vukasin. You became my second family while I was in Geneva. I'll miss the skiing, trips into the city, games, and general banter.

To my undergraduate cohort from Exeter. Though our meet ups are rare, I always look forward to our group holidays and hope they continue far into the future.

Lorenza Iacobuzio, you get a special mention. While we weren't especially close until recently, you've been my sage, partner in food, and exceptional friend when I've needed it the most.

Finally, and most importantly, my family deserve my thanks. You have been there for me since the very beginning supporting me through school, undergrad, PhD, and every other endeavour. To my grandparents, uncles, aunties, and cousins, my siblings Joe, Lydia, Sitara, and Arjen, my stepmum Nadia, and my cat Pixie, I am sincerely grateful to have you in my lives. But above all else, to my mum Meeta and dad Kiron, I could not have done any of this without you and as such, you have my deepest and most heartfelt love and gratitude.

\end{SingleSpace}
\clearpage
